Towards Practical First-Order Model Counting

### 2024-08-12 doc/paper/paper.tex
Minor edits

Since \textsc{FastWFOMC} does not support many-sorted logic, our experiments are
limited to formulas with a single domain. However, many of the counting problems
used in the experimental evaluation of
\textsc{Crane}~\cite{DBLP:conf/kr/DilkasB23} become equivalent (in terms of
generating equivalent integer sequences) when restricted to a single domain.
Additionally, comparing \Cranetwo{} and \textsc{FastWFOMC} on a broader set of
problems is challenging because each problem needs to be represented in two
(quite different) logics: \FO{} and \UFO{}. Nevertheless, the three benchmarks
cover a wide range of possibilities.

Among the three benchmarks, the `friends' benchmark stands out as it uses
multiple predicates and can be expressed in \FO{} using just two variables
without cardinality constraints or counting quantifiers. The `functions'
benchmark, on the other hand, can still be handled by all the algorithms, but it
requires cardinality constraints, counting quantifiers, or more than two
variables. Lastly, the `bijections' benchmark is an example of a formula that
\textsc{FastWFOMC} can handle but \textsc{ForcLift} cannot.

@@

Another aspect of the experimental results that deserves separate discussion is
compilation. Both Julia and Scala use just-in-time (JIT) compilation, which
means that \textsc{FastWFOMC} and \textsc{ForcLift} take longer to run on the
smallest domain size, where most JIT compilation occurs. In the case of
\Cranetwo{}, it is only run once per benchmark, so the JIT compilation time is
included in its overall runtime across all domain sizes. Additionally, while
\textsc{ForcLift}'s compilation is generally faster than that of \Cranetwo{},
neither significantly affects the inference time. Specifically,
\textsc{ForcLift} compilation typically takes around \SI{0.5}{\second}, while
\Cranetwo{} compilation takes around \SI{2.3}{\second}.

>>>>

Since \textsc{FastWFOMC} does not support many-sorted logic, our experiments are
limited to formulas with a single domain. However, many of the counting problems
used in the experimental evaluation of
\textsc{Crane}~\cite{DBLP:conf/kr/DilkasB23} become equivalent (in terms of
generating equivalent integer sequences) when restricted to a single domain.
Additionally, comparing \Cranetwo{} and \textsc{FastWFOMC} on a broader set of
problems is challenging because each problem needs to be represented in two
(quite different) logics: \FO{} and \UFO{}.

Nevertheless, the three benchmarks cover a wide range of possibilities. The
`friends' benchmark stands out as it uses multiple predicates and can be
expressed in \FO{} using just two variables without cardinality constraints or
counting quantifiers. The `functions' benchmark, on the other hand, can still be
handled by all the algorithms, but it requires cardinality constraints, counting
quantifiers, or more than two variables. Lastly, the `bijections' benchmark is
an example of a formula that \textsc{FastWFOMC} can handle but \textsc{ForcLift}
cannot.

@@

Another aspect of the experimental results that deserves separate discussion is
compilation. Both Julia and Scala use just-in-time (JIT) compilation, which
means that \textsc{FastWFOMC} and \textsc{ForcLift} take longer to run on the
smallest domain size, where most JIT compilation occurs. In the case of
\Cranetwo{}, it is only run once per benchmark, so the JIT compilation time is
included in its overall runtime across all domain sizes. Additionally, while
\textsc{ForcLift}'s compilation is generally faster than that of \Cranetwo{},
neither significantly affects overall runtime. Specifically, \textsc{ForcLift}
compilation typically takes around \SI{0.5}{\second}, while \Cranetwo{}
compilation takes around \SI{2.3}{\second}.
